---
title: "Stats102A, Summer 2023 - Homework 2"
author: "Luke Villanueva - 206039397"
date: "`r format(Sys.Date(), '%D')`"
output:
  pdf_document:
  extra_dependencies: ["amsmath"]
---

```{r include=F}
library(stringr)
library(tidyverse)
library(tidyr)

source("206039397_stats102a_hw2.R")
```

# 1. Teacher's Gradebook

## a.

```
FUNCTION gen_gradebook()

  SET seed to UID
  
  FUNCTION gen_score
    RETURN random double from 0-1 * 100, change to integer
  END FUNCTION
  
  FUNCTION gen_UID
    SET rnum = random double from 0-3
    SET num = character of rnum, substring from index 0-9
    SET indx = index of "." in num
    SET res = substring of num from 1 to indx + substring of num from indx + 1 to length of num
    RETURN res
  END FUNCTION

  SET nStudents = 100
  
  SET gradebook = data frame, row = nStudents, col = 11
  
  SET column names of gradebook = "UID","Homework_1",
                           "Homework_2","Homework_3",
                           "Homework_4","Homework_5",
                           "Quiz_1","Quiz_2","Quiz_3",
                           "Quiz_4","Quiz_5"
                           
  SET uid = empty vector
  
  FOR i in vector 1:nStudents
    SET uid = uid + gen_UID()
  END FOR
  SET gradebook[,"UID"] = uid
  
  FOR col in columns of gradebook WITHOUT UID
    SET grades = empty vector
    FOR i in 1:nStudents
      SET grades = grades + gen_score()
    END FOR
    SET gradebook[,col] = grades
  END FOR

  RETURN gradebook

END FUNCTION
```

```{r}
gradebook <- gen_gradebook()
```

## b.

```

FUNCTION na.replace(vec, percent = 10)

  FUNCTION get_random_index(vec)
    RETURN random number from 1-length of vec INCLUSIVE
  END FUNCTION
  
  SET res = vec
  
  IF percent is not number
    STOP
  END IF
  IF any NA's in vec
    WARNING
  END IF
  IF percent % 1 != 0
    STOP
  END IF
  IF percent < 0
    STOP
  END IF
  IF percent == 0
    STOP
  END IF
  IF percent > 100
    STOP
  END IF
  
  SET percent = percent / 100
  
  SET needed_nas = roundup(percent * length(vec))
  
  SET n_na = length(vector of indices of na's in vec)
  
  SET na_left = needed_nas - n_na
  
  WHILE na_left > 0
  
    SET indx = NA
    
    WHILE indx is NA
      SET indx = get_random_index(vec)
    END WHILE 
    
    SET res[indx] = NA
    SET na_left = na_left - 1
    
  END WHILE
  
  RETURN res

END FUNCTION

```

```{r na.replace implementation}

# randomly replace data in vector with NA
# percent is inputted as an integer
# function will AT LEAST make the NA percent, but it can over do it
na.replace <- function(vec,percent = 10)
{
  
  # function to pick random index in vec
  get_random_index <- function(vec)
  {
    floor(runif(1,min = 1, max = length(vec)+1))
  }
  
  # result vector
  res <- vec
  
  # default of percent is 10
  
  # input validation for percent
  if(!is.numeric(percent))
  {
    stop("inputted percent is not numeric")
  }
  if(length(percent) != 1)
  {
    stop("inputted percent is a vector, and should be one value")
  }
  
  # input validation for vec
  # send warning if there are na's in vec already, telling user
  # that those na's will count too
  if(any(is.na(vec)))
  {warning("there exists NA's in inputted vector. 
           These NA's will count towards the percentage of NA's")}
  
  # check if percent is inputted as a decimal
  if(percent %% 1 != 0)
  {stop("inputted percent is decimal, but should be integer")}
  
  # check if percent is negative
  if(percent < 0)
  {stop("inputted percent should be nonnegative")}
  
  # check if percent is 0
  if(percent == 0)
  {warning("inputted percent is 0, should be positive; function will still run")}
  
  # check if percent is bigger than 100 percent
  if(percent > 100)
  {stop("inputted percent is greater than 100, should be from 1-100")}
  
  # get how many to change to reach at least 10 percent
  percent <- percent / 100
  # round up to get AT LEAST percent
  needed_nas <- ceiling(percent * length(vec))

  # count how many NAs, subtract current NA's from 
  n_na <- length(which(is.na(vec)))
  
  
  # get percentage of NA's left to do
  na_left <- needed_nas - n_na
  
  # while there are na left to make
  while(na_left > 0)
  {
    # get random indx
    indx <- NA
    # check if not already NA
    while(is.na(indx))
    {indx <- get_random_index(vec)}
    
    # once we get here, indx is not NA, so change to NA
    res[indx] <- NA
    
    # update na_left counter
    na_left <- na_left - 1
  }
  
  # no more needed na's, so return vec
  return(res)
  
}

```

```{r}

gradebook$Homework_4 <- na.replace(gradebook$Homework_4,10)
gradebook$Quiz_4 <- na.replace(gradebook$Quiz_4,10)

# show there are 10 na's in specific columns
sum(is.na(gradebook$Homework_4))
sum(is.na(gradebook$Quiz_4))

```

## c. Messy Impute - messy_impute()

```

FUNCTION messy_impute()

  INPUT df, center = "mean", margin, ...
  
  IF df is not the right class
    STOP
  END IF
  
  IF df cols != "UID","Homework_1",
                       "Homework_2","Homework_3",
                       "Homework_4","Homework_5",
                       "Quiz_1","Quiz_2","Quiz_3",
                       "Quiz_4","Quiz_5"
    STOP
  END IF
  
  IF df does not have valid dimensions
    STOP
  END IF
  
  IF center is not character OR chars in center > 1 OR center != "mean" or "median"
    STOP
  END IF
  
  IF margin is not numeric OR length margin > 1
    STOP
  END IF
  
  
  FOR col in columns of df[1:length(df)]
    SET df[,col] = numeric of df[,col]
  END FOR
  
  SET fun = NULL
  IF center == "mean"
    SET fun = mean()
  END IF
  ELSE
    SET fun == median()
  END ELSE
  
  
  IF margin == 1
  
    FOR row in rows of df
      IF any of df[row,] is NA
        SET impute_val = round(fun(double(df[row,-1]), remove NA's, ...), digitRound = 2)
        
        FOR i in indices of df[row,] == NA
          SET df[row,i] == impute_val
        END FOR
        
    END FOR
  
  END IF
  
  ELSE 
  
    FOR col in cols of df
      IF any of df[,col] is NA
        SET impute_val = round(fun(double(df[,col]), remove NA's, ...), digitRound = 2)
        
        FOR i in indices of df[,col] == NA
          SET df[i,col] == impute_val
        END FOR
    END FOR
  
  END ELSE

  FOR col in cols of df
    SET df[i,col] = formatted to character rounded to 2 digits
  END FOR

  RETURN df   

END FUNCTION

```

## d. 

```{r}

# selecting and printing two students with missing HW 4 and missing Q4

print("Students missing Homework_4:")

print(gradebook %>% filter(is.na(Homework_4)) %>% slice(1:2))

print("Students missing Quiz_4:")

print(gradebook %>% filter(is.na(Quiz_4)) %>% slice(1:2))

```

```{r}
# get UID of example students
na_uids <- gradebook %>% mutate(index = rownames(gradebook)) %>% filter(is.na(Homework_4)) %>% slice(1:2)

na_uids <- rbind(na_uids, gradebook %>% mutate(index = rownames(gradebook)) %>% filter(is.na(Quiz_4)) %>% slice(1:2))

# get indices of example students
indx <- as.numeric(na_uids %>% pull(index))

# print example students
gradebook %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)

```

```{r}
# messy_impute cases
messy_impute(gradebook,"mean", 1) %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)
messy_impute(gradebook,"mean", 2) %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)
messy_impute(gradebook,"median", 1) %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)
messy_impute(gradebook,"median", 2) %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)
messy_impute(gradebook,"mean", 1, trim = 0.25) %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)
messy_impute(gradebook,"mean", 2, trim= 0.25) %>% mutate(index = as.numeric(rownames(gradebook))) %>% slice(indx) %>% select(-index)

```

## e. gradebook_tidy - tidify()

```{r}

# assumes there is gradebook name
tidify_gradebook <- function()
{
  
  # INPUT VALIDATION
  if(!exists("gradebook"))
  {stop("there should exists a \"gradebook\" object that holds UID, Homeworks, Quizzes")}
  
  # work with homework, mutate order of homework
  hw <- gradebook %>% select(c("UID","Homework_1","Homework_2","Homework_3","Homework_4","Homework_5")) %>% pivot_longer(c("Homework_1","Homework_2","Homework_3","Homework_4","Homework_5"), names_to = "Homework_Num", values_to = "HW_Score") %>% mutate(order_num = as.numeric(substr(Homework_Num,10,length(Homework_Num))))
  
  # work with quizzes, mutate order
  qz <- gradebook %>% select(c("UID","Quiz_1","Quiz_2","Quiz_3","Quiz_4","Quiz_5")) %>% pivot_longer(c("Quiz_1","Quiz_2","Quiz_3","Quiz_4","Quiz_5"), names_to = "Quiz_Num", values_to = "Q_Score") %>% mutate(order_num = as.numeric(substr(Quiz_Num,6,length(Quiz_Num))))
  
  res <- hw %>% left_join(qz, by = c("UID","order_num")) %>% select(c("UID", "order_num","HW_Score", "Q_Score")) %>% rename(Assgn_Num = order_num, Homework = HW_Score, Quiz = Q_Score)
  
  return(res)
  
}

gradebook_tidy <- tidify_gradebook()

```

## f. tidy_impute()

```

FUNCTION tidy_impute()

  INPUT df, center = "mean", margin, ...
  
  IF df is not the right class
    STOP
  END IF
  
  IF df cols != "UID","Assgn_Number", "Homework", "Quiz"
    STOP
  END IF
  
  IF df does not have valid dimensions
    STOP
  END IF
  
  IF center is not character OR chars in center > 1 OR center != "mean" or "median"
    STOP
  END IF
  
  IF margin is not numeric OR length margin > 1
    STOP
  END IF
  
  
  FOR col in columns of df[3:4]
    SET df[,col] = numeric of df[,col]
  END FOR
  
  SET fun = NULL
  IF center == "mean"
    SET fun = mean()
  END IF
  ELSE
    SET fun == median()
  END ELSE
  
  
  IF margin == 1
  
    FOR row in rows of df
      IF any of df[row,] is NA
        SET impute_val = round(fun(double(df[row,3:4]), remove NA's, ...), digitRound = 2)
        
        FOR i in indices of df[row,] == NA
          SET df[row,i] == impute_val
        END FOR
        
    END FOR
  
  END IF
  
  ELSE 
  
    FOR col in cols of df[3:4]
      IF any of df[,col] is NA
        SET impute_val = round(fun(double(df[,col]), remove NA's, ...), digitRound = 2)
        
        FOR i in indices of df[,col] == NA
          SET df[i,col] == impute_val
        END FOR
    END FOR
  
  END ELSE

  FOR col in cols of df [3:4]
    SET df[i,col] = formatted to character rounded to 2 digits
  END FOR

  RETURN df   

END FUNCTION

```

## g. tidy_impute Demo

```{r}

# get UID of example students
na_uids <- gradebook_tidy %>% mutate(index = rownames(gradebook_tidy)) %>% filter(is.na(Homework)) %>% slice(1:2)

na_uids <- rbind(na_uids, gradebook_tidy %>% mutate(index = rownames(gradebook_tidy)) %>% filter(is.na(Quiz)) %>% slice(1:2))

# get indices of example students
indx <- as.numeric(na_uids %>% pull(index))

# print example students
gradebook_tidy %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)

```

```{r}


tidy_impute(gradebook_tidy,"mean", 1) %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)


tidy_impute(gradebook_tidy,"mean", 2) %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)

tidy_impute(gradebook_tidy, "median", 1) %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)

tidy_impute(gradebook_tidy, "median", 2) %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)

tidy_impute(gradebook_tidy, "mean", 1, trim = 0.25) %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)

tidy_impute(gradebook_tidy, "mean", 2, trim = 0.25) %>% mutate(index = as.numeric(rownames(gradebook_tidy))) %>% slice(indx) %>% select(-index)


```


# 2. Short Answers

## a.

```{r}

gdp <- read.csv("gdp-countries.csv")

# display 10 observations
sampGdp <- gdp[1:10,]
# first 10 cols
print(sampGdp[,1:10])

web <- read.csv("Most Popular websites.csv")

sampWeb <- web[1:10,]
# first 10 cols
print(sampWeb[,1:10])

watch <- read.csv("swiss watch brands.csv")

sampWatch <- watch[1:10,]
# first 10 cols
print(sampWatch[,1:10])

```

The dataset `gdp` displays each country's GDP over time. Only the country's name and GDP were explicitly collected. The row order is assumed to be the years increasing. The rows are GDP over time and the columns are country names.

[LINK TO GDP](https://www.kaggle.com/datasets/iamtushara/gdp-timeseries-data-for-various-countries?select=gdp-countries.csv)

The dataset `web` shows the popularity of each listed website. Website name, where the image of the company can be located, and the years with the respective website popularity were collected. The rows are different websites and the columns are the website names, image location, and years.

[LINK TO WEB](https://www.kaggle.com/datasets/mathurinache/mostpopularwebsites)

The dataset `watch` shows an undisclosed measured metric of different watch brands over time. The rows are different brands and the columns are the metric changing over time.

[LINK TO WATCH](https://www.kaggle.com/datasets/mathurinache/swisswatchbrands)

## b.

### Tidy gdp

```{r}
# print first 10 rows
print("sampGdp First 10 Rows:")
print((sampGdp %>% mutate(time = as.numeric(rownames(sampGdp))) %>% pivot_longer(colnames(sampGdp), names_to = "Countries", values_to = "GDP"))[1:10,])
```

### Tidy web

```{r}
# print first 10 rows
print((web %>% pivot_longer(colnames(web)[-c(1,2)], names_to = "Years", values_to = "Popularity") %>% mutate(Years = substr(Years,2,length(Years))))[1:10,]) 
```

### Tidy watch

```{r}
# print first 10 rows
print((sampWatch %>% pivot_longer(colnames(sampWatch)[-c(1,2)], names_to = "Years", values_to = "Measure") %>% mutate(Years = substr(Years,2,length(Years))))[1:10,])
```