---
title: "Stats 20, W23 -- Homework 5"
author: "Luke Villanueva -- 206039397"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

# Question 1

```{r}

char <- c("4", "2", "1", "0")

num <- 0:3

charnum <- data.frame(char, num, stringsAsFactors = TRUE)

```

## a. 

```{r}

as.numeric(char)
as.numeric(charnum$char)

typeof(char)
typeof(charnum$char)

char
charnum$char

as.numeric(factor(c("10","5","1"), ordered = TRUE))

```

Because `char` is itself just a character vector, the type casting can take the raw values of the strings and turn them into numbers. However, `charnum$char` returns the numbers as factored data. This turns the column into a factor, which is casted into their ordering indices by the factor order.

## b.

```{r}
as.numeric(as.character(charnum$char))
```

# Question 2

```{r}
simple_list <- list("vector" = 1:10, "matrix" = matrix(6:1, nrow = 3, ncol = 2))

simple_list$NULL <- NULL
```

## a.

1. NULL is a keyword in R. So, making any object or function just named NULL will not work, for it is a reserved word. 

2. Even if it did successfully create a column named NULL, the value of that column is NULL, so appending a value to a nonexistent column will definitely throw out an error of a nonexistent initialized object.

## b.

```{r}

simple_list$null <- list(NULL)

length(simple_list) == 3

```

## c. 

```{r}

simple_list$vector <- list(NULL)

```

# Question 3

```{r}

my_scale <- function(x)
{
  
  a <- 0
  b <- 1
  
  # check if there exists attributes to x
  # if there exists index for both, a and b exist 
  if(length(which(names(attributes(x)) == "a")) != 0 & length(which(names(attributes(x)) == "b")) != 0)
  {
    a <- attributes(x)$a
    b <- attributes(x)$b
  }
  
  # else, just continue the algorithm
  a + (((x - min(x)) * (b - a)) / (max(x) - min(x)))
  
}

# test
x <- 1:20

# test without a and b
my_scale(x)
my_scale(my_scale(x))

# add a and b
attributes(x)$a <- 5
attributes(x)$b <- 2

# test
my_scale(x)
my_scale(my_scale(x))

```

# Question 4

## a. 





