---
title: "Stats 20, W23 -- Homework 3"
author: "Luke Villanueva -- 206039397"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

## Question 1

### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 1a}

# store values in vector
commute_times <- c(14,12,20,19,15,20,28,20,20,18)

# boundary values based on 1 std dev
upper <- mean(commute_times) + sd(commute_times)

lower <- mean(commute_times) - sd(commute_times)

# to get which days, make seq indexing the vector
days <- seq_along(commute_times)

# logical index: get indices where |mean - std dev| > 1 or
days_outside_sd <- days[commute_times < lower | commute_times > upper]

# prints out days where time greater than 1 std dev
days_outside_sd

# prints out times on those days
commute_times[days_outside_sd]

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL) -->

The days (or indices) where the days are past 1 standard deviation are days 1, 2, and 7. In addition, their values are 14, 12, and 28, respectively.

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 1b}

# do contra positive of logical index, get vals in boundaries
days_inside_sd <- days[-days_outside_sd]

# days
days_inside_sd

# and their respective values
commute_times[days_inside_sd]

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL) -->

The days that stayed within 1 standard deviation are days 3, 4, 5, 6, 8, 9, 10. Their values are 20, 19, 15, 20, 20, and 18, respectively.

### (c)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 1c}

num_days_in_sd <- length(days) - length(days_outside_sd)

# print out num_days_in_sd / total days

num_days_in_sd / length(days)

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL) -->

The proportion of Andy's commute time within 1 standard deviation is 0.7 (i.e. 70% of days were within the standard deviation).

## Question 2

```{r Question 2}
NA & TRUE
NA & FALSE
NA | TRUE
NA | FALSE
```
<!-- ENTER YOUR EXPLANATION BELOW, FEEL FREE TO EDIT THE ABOVE CODE CHUNK -->

For the `&` operator, the compiler does not have to check both sides if and only if one side is already false. The entire statement will automatically be false if one expression is false. 

And so, this supports why `NA & TRUE` is NA. Since the right side is true, the compiler has to check what the left side is. Since it's NA, the output will also be NA.

The reason why `NA & FALSE` is false is because the right expression is already false. So, there is no reason for the compiler to check the left side, for the result will be false regardless.

Now, for the `|` operator, the compiler does not have to check both sides if and only if one expression is true. If one side is already true, there is no reason to check the other side.

And so, this supports the output of `NA | TRUE`. Since one side is already true, there is no reason to check the other side. So, the output is `TRUE`

On the other hand, since `NA | FALSE` has false on the right side, the compiler has to check what `NA` results in, but it's NA. So, the result is NA.

## Question 3

<!-- PUT YOUR get_minimum_coins() FROM HW 1 HERE -->
```{r Question 3 background} 

get_minimum_coins <- function(cents)
{

  # declare coin vals
  q_val <- 25
  d_val <- 10
  n_val <- 5
  p_val <- 1
  
  # declare coin amts based on floor division
  q <- cents %/% q_val
  
  # update cents amt
  cents <- cents - (q * q_val)
  
  # Note: repeat process for the other coins
  
  d <- cents %/% d_val
  cents <- cents - (d * d_val)
  
  n <- cents %/% n_val
  cents <- cents - (n * n_val)
  
  p <- cents %/% p_val
  cents <- cents - (p * p_val)
  
  # total up coin amounts
  total_coin <- q + d + n + p
  
  # print coin amt
  total_coin
}

```

### (a)
<!-- ENTER YOUR ANSWER BELOW -->

<!-- USE THE CODE CHUNK BELOW IF YOU'D LIKE -->
```{r Question 3a}

```

### (b)
<!-- ENTER YOUR ANSWER BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 3b}

```

## Question 4

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 4a}

```

###
<!-- USE THE CODE CHUNK BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 4b}

```


## Question 5

### (a)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5a}

```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5b}

```

### (c)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 5c}

```
<!-- ENTER YOUR EXPLANATION BELOW (USING INLINE R AS YOU WILL). ONE SENTENCE OR 'MY ANSWER IS __' IS FINE. -->

## Question 6
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 6}

```

## Question 7

### (a)
<!-- CREATE THE FUNCTION IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 7a}

```

### (b)
<!-- USE THE CODE CHUNK BELOW -->
```{r Question 7b}

```

### (c)
<!-- CREATE THE FUNCTIONs IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTIONs AND THEIR INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 7c}

```
<!-- REMINDER TO USE my_ifelse() TO WRITE THE FUNCTIONS. THERE ARE OTHER WAYS TO WRITE THESE FUNCTIONS WITHOUT USING my_ifelse() BUT PLEASE FOLLOW THE DIRECTIONS. -->

## Question 8

### (a)
<!-- CREATE THE FUNCTIONs IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTIONs AND THEIR INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 8a}

```

### (b)
<!-- USE THE CODE CHUNK BELOW, IT CAN BE JUST COMMANDs; NO NEED TO STORE OBJECTs -->
```{r Question 8b}

```

## Question 9
<!-- MAKE SURE dna.RData IS IN THE SAME DIRECTORY (FOLDER) AS THIS .RMD -->
```{r Question 9 background}
load("dna.RData")
```

### (a)
<!-- CREATE THE FUNCTION AND TEST IT AS ASKED IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 9a}

```

### (b)
<!-- CREATE THE FUNCTION AND TEST IT AS ASKED IN THE CODE CHUNK BELOW -->
<!-- NAME THE FUNCTION AND ITS INPUTS EXACTLY AS STATED IN THE ASSIGNENT -->
<!-- PLAN OUT YOUR APPROACH BEFORE YOU START CODING ANY FUNCTION. -->
```{r Question 9b}

```
