---
title: "Stats 20, Winter 2023 - Homework 1"
author: "Luke Villanueva - 206039397"
date: "`r format(Sys.Date(), '%D')`"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

<!--Question 1-->
## Question 1

<!--a.-->
### a.

“**Sometimes you have to make the hardest climb to see the most beautiful sunrise.** I read that
once on an old lady’s decorative pillow, but it is really how I feel today. I’ve climbed a very weird and rocky
mountain, and it was a pain in the ass, and my legs are tired, and I’m starving. But *the sun is rising over a
sea of love and waffles and possibility.* So *I’m just gonna relax and take a deep breath and enjoy this view for
as long as I possibly can.*” – Leslie Knope (from Parks and Recreation)

<!--b.-->
### b.

| Homework | Midterm | Final Exam | Final Project |
|---------:|--------:|-----------:|--------------:|
| 18% | 30% | 40% | 12% |

<!--c.-->
### c. 

- 2022
  - Fall
    - MATH115A - Linear Algebra
    - LBR STD 10 - Introduction to Labor and Workplace Studies
    - MATH42 - Introduction to Data-Driven Mathematical Modeling: Life, Universe, and Everything
  - Winter
     - MATH131A - Real Analysis
     - STATS20 - Introduction to Statistical Programming with R
     - STATS100A - Introduction to Probability
     
<!--d.-->
### d. 

![Source: [Kitten Link](https://www.google.com/search?q=kitten+picture&client=opera-gx&hs=a4a&sxsrf=AJOqlzUyTpakAFlj16hyoNA8lsB0O7yu8Q:1673824800437&source=lnms&tbm=isch&sa=X&ved=2ahUKEwjXv5Wl28r8AhVAL0QIHerkBy8Q_AUoAXoECAEQAw&biw=1879&bih=939&dpr=1#imgrc=NdjwHMGdfG55NM)](kitten.jpg)


<!--Question 2-->
## Question 2

<!--a.-->
### a.

I think these videos are showing a lack of clear and concise communication between two parties. Therefore, the result is not exactly what either party desired in the first place. And so, I think this is related to the class because the professor and the student want to work together to accomplish a task (i.e. learning a certain statistic/R programming concept or complete a project). However, if either side provides vague and misleading information, the result will not be satisfactory. So, it is very important in a class that's introducing a new concept to students for communication to be concise and clear-cut. This way, the professor, graders, and students will all be successful.

<!--b.-->
### b. 

The following recipe is my own version of grilled cheese:

1. Obtain the following materials:
  - frying pan
  - spatula
  - 1 stick of butter 
  - 1 loaf of sliced bread
  - 1 pack of prepackaged sliced cheese
2. Turn the stove to medium-high heat
3. Wait for 2-3 minutes for the pan to heat up, while doing so, open the packaged loaf of bread
4. Take out 2 slices of bread and lay them on a flat surface
5. Open the packaged cheese and take out 2 pieces of cheese and lay them next to the bread
6. Open one end of the stick of butter and rip off the packaging so that 1 inch of the stick of butter is exposed
7. With the pan heated, grab the unexposed part of the butter stick and press the exposed end against the pan, making circle motions in order to cover the pan with melted butter (there should be sizzling and the butter should be a bit browned)
8. Grab one slice of bread on the flat surface and press one side of the bread into the pan, letting the butter seep into the bread
9. Open 1 slice of the packaged cheese and lay the slice on the bread in the pan
10. Let the bread sear while occasionally grabbing the pan and tilting it to let the melted butter surround the bread more
11. In the same pan, melt more butter the same way as described before
12. Lay the next slice of bread in the pan, and open the second slice of cheese and lay it onto that new slice of bread
13. Use the spatula to flip the first slice of bread onto the second slice, so that the sides with cheese are facing each other
14. Continue cooking the sandwich until both sides are golden brown or crispy

<!---Question 3-->
## Question 3

<!--a.-->
### a. 

<!--i.-->
(i) 

Calculate $\frac{6^8}{16^5}$

```{r}
( 6^8 ) / ( 16^5 )
```

<!--ii.-->
(ii)

Calculate $101^{- \frac{8}{7}}$

```{r}
101^(-8/7)
```

<!--iii.-->
(iii)

Calculate $8 + 3 * \pi - 3 + 4/2 * -1$

```{r}
8 + ( 3 * pi ) - 3 + (4 / 2) * (-1)
```

<!--b.-->
### b.

<!--i.-->
(i.)

```{r}
# log base 2 on 54
log(54,2)
```

<!--ii.-->
(ii.)

```{r}
exp(1)^( exp(1)^( sqrt(2) )  )
```

<!--iii.-->
(iii.)

```{r}
( (exp(1)^(pi)) + (exp(1)^(-pi)) ) / ( 2 )
```

<!--c.-->
### c.

<!--i.-->
(i.)

To find number of times 25 goes into 119, we can use division and ignore the decimal.

```{r}
119 / 25
```

So, we know that 25 can go into 119 by 4 times.

<!--ii.-->
(ii.)

The remainder can be found by using modulus.

```{r}
119 %% 25
```

So, we know that the remainder after 25 goes into 119 by 4 times is 19.



<!--Question 4-->
## Question 4

<!--a.-->
### a.

<!--i.-->
(i)

```{r}

vol_1 <- (4 / 3) * pi * (1^3)
vol_1

```

<!--ii.-->
(ii)

```{r}

vol_4 <- (4 / 3) * pi * (4^3)
vol_4

```

<!--b.-->
### b. 

```{r}

vol_sphere <- function(r = 1)
{
  
  vol <- (4 / 3) * pi * (r^3)
  vol
  
}

```

<!--c.-->
### c. 

<!--i.-->
(i.)

```{r}

vol_sphere()
vol_1

```

The two values are the same, so default value of 1 in `vol_sphere()` is valid.

<!--ii.-->
(ii.)

```{r}

vol_sphere(4)
vol_4

```

The two values are the same, so value of `vol_sphere()` can change depending on inputted parameter.



<!--Question 5-->
## Question 5

<!--a.-->
### a.

```{r}

# Param: x = num of successes,
# n = num of trials, 
# p0 = hypothesized pop proportion
z_prop <- function(x,n,p0)
{
  
  # convert input successes into proportion
  x <- x/n

  z <- ( x - p0 ) / sqrt( (p0*(1-p0)) / (n) )
  
  z
}

```

$\hat{p}$ is the value of x.

<!--b.-->
### b.

```{r}

z_prop(10,13,0.6)

```

<!--c.-->
### c. 

This z-statistic is the number of standard deviations from the mean value from the ideal population. Depending on the significance value, the z-statistic can lead to the null hypothesis being rejected or accepted. The null hypothesis in this case would be that 60% of Pawnee's residents like waffles over pancakes. This would call for a two-tailed test.

<!--d.-->
### d.

```{r}

z_prop(30, 39, 0.6)

```

The z-statistic changed because the proportion of the sample that likes waffles over pancakes changed in this new sample. This can be utilized to accept or reject the fact that the supposed hypothesis is that 60% of Pawnee's residents likes waffles over pancakes.

<!--Question 6-->
## Question 6

<!--a.-->
### a.

```{r}

lease_calc <- function( msrp, price, down, n = 36, res = 0.6, mf = 0.001, tax = 0.095 )
{
  
  capitalized_cost <- price - down
  
  residual_value <- msrp * res
  
  monthly_depreciation <- ( capitalized_cost - residual_value ) / n
  
  monthly_finance_charge <- ( capitalized_cost + residual_value ) * mf
  
  sub_total <- monthly_depreciation + monthly_finance_charge
  
  total <- sub_total * ( 1 + tax  )
  
  total
  
}

```

<!--b.-->
### b. 

```{r}

lease_calc(31495,29895,2500, 36, 0.52, 0.0016, 0.095)

```

<!--Question 7-->
## Question 7

<!--a.-->
### a.

To get the minimum number of coins, we would get the most coins starting from the bigger denominations.

So, we can fit 1 quarter into 47 cents. That leaves 22 cents. We can fit 2 dimes. That leaves 2 cents. And lastly, 2 pennies.

So, the least amount of coins would be 5 coins: 1 quarter, 2 dimes, 2 pennies.

<!--b.-->
### b.

For an arbitrary number of cents, the generalized logic would be similar to the previous part:

1. Starting from the biggest denomination, fit the coin until it cannot fit as a whole.
2. Move down to the next denomination and fill as much as you can with that coin.
3. Continue this process until the total cent amount is covered.

The following is the code logic for the process:

```

INPUT cents

LET q_val = 25, d_val = 10, n_val = 5, p_val = 1

LET q = 0, d = 0, n = 0, p = 0

SET q = cents FLOOR q_val
SET cents = cents - (q * q_val)

SET d = cents FLOOR d_val
SET cents = cents - (d * d_val)


SET n = cents FLOOR n_val
SET cents = cents - (n * n_val)

SET p = cents FLOOR p_val
SET cents = cents - (p * p_val)

```

<!--c.-->
### c.

```{r}

get_minimum_coins <- function(cents)
{

  # declare coin vals
  q_val <- 25
  d_val <- 10
  n_val <- 5
  p_val <- 1
  
  # declare coin amts based on floor division
  q <- cents %/% q_val
  
  # update cents amt
  cents <- cents - (q * q_val)
  
  # Note: repeat process for the other coins
  
  d <- cents %/% d_val
  cents <- cents - (d * d_val)
  
  n <- cents %/% n_val
  cents <- cents - (n * n_val)
  
  p <- cents %/% p_val
  cents <- cents - (p * p_val)
  
  # total up coin amounts
  total_coin <- q + d + n + p
  
  # print coin amt
  total_coin
}

```

<!--d.-->
### d.

<!--i.-->
(i.)

Testing `get_minimum_coins()` with 31 cents:

```{r}

get_minimum_coins(31)

```

<!--ii.-->

(ii.)

Testing 48 cents:

```{r}

get_minimum_coins(48)

```

<!--iii.-->
(iii.)

```{r}

get_minimum_coins(139)

```

<!--e.-->
### e.

To max the coins, we have to pick a number that would maximize each denomination. Unfortunately, we can't do this efficiently because we are limited to the given function `get_minimum_coins()`, which considers quarters first and descends onward from there. However, we can somewhat start with pennies first by focusing on the rightmost digit.

Initially, the number has to have a 4 or 9 in the rightmost digit place. If it were 5 or 0, the total can be replaced with a nickel or dime, so to maximize the amount of pennies, the total's rightmost digit must be a 4 or 9. Thus, the max amount of pennies would be 4.

The function will first consider quarters as the initial denomination. So, the max quarters allowed would be 3 since 4 quarters is exactly 100 cents. So, we have 75 cents of quarters plus 4 pennies, which equals to the total so far of 79 cents.

Next, the max amount of dimes would be found by the function. The most amount of dimes that can be added to 79 cents is 2 because 2 dimes is 20 cents. 79 plus 20 equals 99. So, 99 is our total cents.

Since 99 is 1 less than 100, there is no more room to grow in amount, so the max amount of coins is 3 quarters, 2 dimes, and 4 pennies. This totals up to 9 coins.

To verify:

```{r}

get_minimum_coins(99)

```


<!--Question 8-->
## Question 8

<!--a.-->
### a.

`exp1()` is expected to provide a different output than `exp2()` because `exp2()` has a default value for its parameter. So, even if the two functions have the same body code, the values for `x` are different. 

<!--b.-->
### b.

By placing a function within a function with two uninitialized parameters, we are creating a function that sets up the code for another function. Meaning, the outer most `function()` code block gets used first, placing the parameter into the given code block and then assigning the function code block to the variable name (which is `exp_y` in this case). Afterward, the code can be inputted with another parameter because the function has been "set up" already.


<!--c.-->
### c.

```{r}
x <- 3
y <- 4

exp_y <- function(x)
{
  
  function(y)
  {
    
    x^y
    
  }
  
}

# to output 15^4, input 15 first and then 4
(exp_y(15))(4)

```