---
title: "Stats101A, Spring 2023 - Homework 1"
author: "Luke Villanueva - 206039397"
date: "`r format(Sys.Date(), '%D')`"
output: 
  pdf_document: default
---

# Problem 1

## a. 

confidence interval = [mean - margin of error, mean + margin of error]

margin of error = t crit val of confidence * standard of error

standard of error = standard deviation / sqrt(n)

And so, to find the interval, 

```{r}

# standard error
SE <- 24757 / sqrt(30)

# margin of error
MOE <- abs(qt(0.05, df = 29)) * SE

# vector of the 95% confidence interval
c(23606 - MOE, 23606 + MOE)

```

So, the 95% confidence interval is (15925.96, 31286.04). 

## b. 

Because only the sample standard deviation was known, the t value was taken instead of the the z value.

## c. 

Yes, this is because the samples come from the same population and the sample size is the same too. The definition of the confidence interval means that 95% of all samples' means will fall within that interval. And so, this implies that the second average's mean will have a 95% chance of landing in the found confidence interval.

## d. 

Two tailed test:

Null hypothesis: The mean income in the US in the year 2000 is $25,000.

Alternative hypothesis: The mean income in the US in the year 2000 is not $25,000.

Doing a t-test, we need to calculate the t score, get the critical value for 5% significance, and reject or fail to reject null hypothesis:

```{r}

# t score
t_score <- abs((25000 - 23606) / (24757 / sqrt(30)))
t_score

# t crit val 5% sig, two tailed, divide by 2
crit <- abs(qt(0.025, df = 29))
crit 

# should we reject null hypothesis?
t_score > crit

```

The sample data does not provide enough information to disprove the claim that the mean is $25,000.

## e.

```{r}

# brute force, find the t score
# checking if t_score can be rejected
t_score > abs(qt(0.38, df = 29))

# since two tailed test, multiply by 2
2*0.38

```
This means the smallest significance value would be 76% significance, which is really big and does not help with pointing out any statistically significant results.

# 2. 


